---
title: "Data Manipulation"
output: github_document
---

```{r setup}
#option command i to insert a code chunk
library(tidyverse)
```

## Load in the FAS Litters data
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

Choose some columns and not others.

```{r}
select(litters_df, group, gd0_weight)
```

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

Renaming columns...

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

Select helpers

```{r}
#select helper functions
select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, litter_number, pups_survive, everything()) #moves litter number and pups survive to beginning
```

```{r}
relocate(litters_df, litter_number, pups_survive) #moves litter number and pups survive to beginning
```

## 'filter'

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```


```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Con7")
```

```{r}
filter(litters_df, group %in% c("Con7","Mod8")) #where group is either of these two options
```

You will often filter using comparison operators (>, >=, <, <=, ==, and !=). You may also use %in% to detect if values appear in a set, and is.na() to find missing values. The results of comparisons are logical – the statement is TRUE or FALSE depending on the values you compare – and can be combined with other comparisons using the logical operators & and |, or negated using !.

Some ways you might filter the litters data are:

* gd_of_birth == 20
* pups_born_alive >= 2
* pups_survive != 4
* !(pups_survive == 4) #not equal to 4
* group %in% c("Con7", "Con8")
* group == "Con7" & gd_of_birth == 20
A very common filtering step requires you to omit missing observations. You can do this with filter, but I recommend using drop_na from the tidyr package:

* drop_na(litters_data) will remove any row with a missing value
* drop_na(litters_data, wt_increase) will remove rows for which wt_increase is missing.

