---
title: "Data Manipulation"
output: github_document
---

```{r setup}
#option command i to insert a code chunk
library(tidyverse)
```

## Load in the FAS Litters data
```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```

## 'select'

Choose some columns and not others.

```{r}
select(litters_df, group, gd0_weight)
```

```{r}
select(litters_df, group, gd0_weight:gd_of_birth)
```

```{r}
select(litters_df, -litter_number)
```

Renaming columns...

```{r}
select(litters_df, GROUP = group, LITTer_NUmBer = litter_number)
```

```{r}
rename(litters_df, GROUP = group, LiTtEr_NuMbEr = litter_number)
```

Select helpers

```{r}
#select helper functions
select(litters_df, starts_with("gd"))
```

```{r}
select(litters_df, litter_number, pups_survive, everything()) #moves litter number and pups survive to beginning
```

```{r}
relocate(litters_df, litter_number, pups_survive) #moves litter number and pups survive to beginning
```

## 'filter'

```{r}
filter(litters_df, gd0_weight < 22)
```

```{r}
filter(litters_df, gd0_weight >= 22)
```


```{r}
filter(litters_df, gd0_weight >= 22, gd_of_birth == 20)
```

```{r}
filter(litters_df, group == "Con7")
```

```{r}
filter(litters_df, group %in% c("Con7","Mod8")) #where group is either of these two options
```

You will often filter using comparison operators (>, >=, <, <=, ==, and !=). You may also use %in% to detect if values appear in a set, and is.na() to find missing values. The results of comparisons are logical – the statement is TRUE or FALSE depending on the values you compare – and can be combined with other comparisons using the logical operators & and |, or negated using !.

Some ways you might filter the litters data are:

* gd_of_birth == 20
* pups_born_alive >= 2
* pups_survive != 4
* !(pups_survive == 4) #not equal to 4
* group %in% c("Con7", "Con8")
* group == "Con7" & gd_of_birth == 20
A very common filtering step requires you to omit missing observations. You can do this with filter, but I recommend using drop_na from the tidyr package:

* drop_na(litters_data) will remove any row with a missing value
* drop_na(litters_data, wt_increase) will remove rows for which wt_increase is missing.

## 'mutate'

```{r}
mutate(
  litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group)
)
```

## 'arrange'

```{r}
head(arrange(litters_df, group, pups_born_alive), 10)
```

## '%>%'

```{r}
litters_data_raw = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data_clean_names = janitor::clean_names(litters_data_raw)
litters_data_selected_cols = select(litters_data_clean_names, -pups_survive)
litters_data_with_vars = 
  mutate(
    litters_data_selected_cols, 
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group))
litters_data_with_vars_without_missing = 
  drop_na(litters_data_with_vars, wt_gain)
litters_data_with_vars_without_missing
```

```{r}
litters_data_clean = 
  drop_na(
    mutate(
      select(
        janitor::clean_names(
          read_csv("./data/FAS_litters.csv", col_types = "ccddiiii")
          ), 
      -pups_survive
      ),
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)
    ),
  wt_gain
  )

litters_data_clean
```

Both confusing. use piping instead (shift command m for pipe operator).

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>% #shift command m for pipe operator
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(wt_gain)

litters_data
#easier to read
```

```{r}
litters_data = 
  read_csv("./data/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names(dat = .) %>%
  select(.data = ., -pups_survive) %>%
  mutate(.data = .,
    wt_gain = gd18_weight - gd0_weight,
    group = str_to_lower(group)) %>% 
  drop_na(data = ., wt_gain)
```

```{r}
litters_data %>%
  lm(wt_gain ~ pups_born_alive, data = .) %>%
  broom::tidy()
```

